An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-24 VLDB2026/main.tex
Reduce 2.2 sections

	\noindent\textbf{\underline{Graph.}}
	Graph-based ANN search algorithms accelerate queries by building graph indexes on datasets. As illustrated in Figure~\ref{fig:graph}, each data point in the dataset corresponds to a point in the graph, and neighboring vertices (e.g., $a$ and $b$) are connected via edges based on their distance $\textit{dist}(a, b)$. In the graph index, each point maintains connections only to its nearest neighbors to preserve query efficiency. For instance, in Figure~\ref{fig:graph}, the gray point $c$ is connected to 5 neighbors $(a, b, d, e, f)$ and can traverse to them via corresponding edges.

	Given a query point $q$, the graph-based ANN search typically follows a greedy search strategy to find the $k$ nearest neighbors. For illustration, consider the case where $k = 1$ (i.e., 1NN): The algorithm initiates the search from an entry point (the gray point $c$). If there exists a neighbor $n$ (the point $f$) of the current point such that $\textit{dist}(n, q) < \textit{dist}(c, q)$, the algorithm updates the current point to $n$. This process is iteratively repeated—always moving to the neighbor closest to $q$—until no neighbor is found that is closer than the current point. The final point reached (the blue point $j$) is returned as the approximate nearest neighbor of $q$.

	\textcolor{violet}{Graph-based indexes are widely adopted in Hybrid ANN search algorithms due to their excellent performance. To support hybrid queries, existing approaches usually augment the original graph structures by explicitly associating attribute or range information with vertices or edges. In attribute filtering, a common approach is to attach attribute metadata to vertices or edges, allowing the traversal process to skip nodes or edges that do not meet attribute conditions of the query (e.g., NHQ, Filtered-DiskANN, ACORN). In range filtering, edges typically store valid attribute intervals, and traversal proceeds only along edges satisfying the range constraints, effectively narrowing down the search space (e.g., SeRF, DSG).}

	\noindent\textbf{\underline{IVF.}}  IVF-based ANN search algorithms improve computational efficiency by partitioning the dataset into clusters and restricting the search to clusters nearest to the query point. Specifically, as shown in Figure~\ref{fig:ivf}, IVF first selects a subset of data points and applies a clustering algorithm (e.g., K-Means) to obtain a set of centroids $(C_0, C_1, ..., C_6)$, each representing a distinct cluster. Every data point (gray dots) is then assigned to the cluster corresponding to its nearest centroid based on the distance metric, thereby forming the index structure.

	During the query process, given a query point $q$, the algorithm computes distances between $q$ and all centroids, selects a small number of the closest centroids ($C_0, C_1, C_2$), and then performs a search only within the associated clusters to identify the approximate nearest neighbors.

	\textcolor{violet}{IVF-based indexes also offer unique advantages in Hybrid ANN search scenarios, such as naturally supporting pre-filtering and requiring significantly lower memory consumption and smaller index sizes compared to graph-based indexes. With minor modifications, IVF indexes achieve attribute and range filtering by directly excluding vectors that do not meet query constraints through pre-filtering, thus skipping irrelevant vectors during intra-cluster distance computations (e.g., Faiss, Milvus). Moreover, clusters can be further partitioned hierarchically based on attributes (hierarchical partitioning), resulting in finer-grained sub-partitions that significantly enhance the efficiency of attribute filtering (e.g., CAPS, PUCK).}

@@

	\textit{\textbf{\textcolor{blue}{Different Hardware Environments.}}} 
%	\textcolor{cyan}{To investigate the impact of different server configurations on the performance of the algorithms, we conducted additional experiments on Server B using the YT-Audio-Real dataset, and the results are shown in Figure~\ref{fig:Different Computer}. Compared with the experimental results on Server A shown in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}f, the relative ranking of the algorithms remains largely unchanged, indicating that the performance of these methods is not significantly affected by differences in CPU configurations. Due to the lower CPU frequency and smaller RAM capacity of Server B compared to Server A, the QPS values are slightly lower across all methods.}
	\textcolor{red}{To investigate the impact of different server configurations on the performance of the algorithms, we conduct additional experiments on a server equipped with dual Intel(R) Xeon(R) Gold 6330 CPUs @ 2.00GHz using the YT-Audio-Real dataset, and the results are shown in Figure~\ref{fig:attribute-cross-platform}. Compared with the experimental results on the original server, the relative ranking of the algorithms remains largely unchanged, indicating that the performance of these methods is not significantly affected by differences in CPU configurations. Due to the lower CPU frequency and smaller RAM capacity, the QPS values are slightly lower across all methods.}

	\textit{\textbf{\textcolor{blue}{Large-scale Datasets.}}} 
	\textcolor{violet}{As shown in the figure~\ref{fig:attribute big dataset}, On the 100M dataset, UNG and NHQ still exhibit excellent performance, demonstrating their good adaptability to large-scale datasets; however, this also comes with significant memory overhead. The clustering-based algorithm Faiss+HQI\_Batch shows moderate performance but has very low memory overhead, which makes it highly attractive for practical applications. We did not present the experimental results for StitchedVamana because we found that its recall consistently remained at a very low level. Its search starting point is set to 0 by default, and when the entry point is far from the target area, it easily gets stuck in a local optimum, making it difficult to navigate to the vicinity of the correct point.}

@@

\textcolor{blue}{
	We conducted experiments on the text2image-mix dataset. This dataset contains both image and text vectors. It is a more complex multimodal dataset. As shown in Figure~\ref{fig:attribute-multimodal}, although the queries are more complex, UNG and NHQ still perform the best. CAPS is the next best.
	Since the dataset mixes two modalities, the ground truth of the query set usually corresponds to the same modality. The nearest neighbors in the index are often from the same modality because they are closer in distance. Therefore, the search results on this dataset are similar to those in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}. UNG and NHQ perform the best. CAPS clusters based on vector similarity. It actually clusters the vectors into the same modality. Then it uses attribute filtering to divide the data further. This allows CAPS to quickly find nearest neighbors that meet the filter conditions.
	In contrast, the text2image dataset shown in Figure~\ref{fig:exp_6_1} is different. The queries and the dataset are from different modalities. The queries are far from the dataset. The ground truths are also scattered and not clustered together. To find the ground truths, the algorithm often needs to traverse more points. This is a greater challenge for the algorithms.
	%	We conducted extensive scalability experiments on the Text2Image-Mix dataset. This dataset contains both image and text vectors. It represents a more complex multimodal query scenario. As shown in Figure 12, the UNG and NHQ algorithms still achieve the best performance, followed closely by CAPS.
	%	In the Text2Image-Mix dataset, the ground truth of the query vector is usually the corresponding vector in the same modality. Since the neighbor relationships in the graph are often based on vector distance, vectors of similar modalities naturally form closer connections in the graph. Therefore, during the search process, the algorithm tends to traverse efficiently within the same modality. This makes the search mechanism similar to our experiments on single-modality datasets, thus maintaining high performance.
	%	The good performance of the CAPS algorithm is due to its unique hierarchical space partitioning strategy. It first clusters vectors based on similarity. In multimodal data, this clustering naturally groups vectors of similar modalities into the same cluster. Then, CAPS further partitions these clusters using attributes. This allows it to quickly find the nearest neighbors that meet the filtering conditions.
	%	It is worth noting that, compared to the OOD dataset discussed in the "Dataset Impact" section, the Text2Image-Mix dataset also involves modality differences. However, its complexity lies in modality mixing. The OOD dataset, on the other hand, has a modality gap. This often requires traversing a broader and more sparse search space to find the ground truth. It poses a more severe challenge to the algorithm.
a}

@@

	\textit{\textbf{\textcolor{blue}{Different Hardware Environments.}}}
	\textcolor{cyan}{Consistent with Section 4.2.4, we conduct additional range filtering experiments on the YT-Audio-Real dataset using the same server configuration, with results presented in Figure~\ref{fig:range-cross-platform}. When compared with the experimental results of the original server shown in Figure~\ref{fig:exp_8_2}a, our findings remain consistent with those reported in Section 4.2.4. For brevity and due to space limitations, we omit further discussion of these redundant results.}
%	\textcolor{cyan}{Consistent with Section 4.2.4, we conducted additional range filtering experiments using the YT-Audio-Real dataset on the same server, with results shown in Figure \ref{fig:Different Computer range}. Compared with the original server's experimental results (Figure \ref{fig:exp_8_2}($e$)), the conclusions remain identical to those described in Section 4.2.4. Due to space constraints, these details are not reiterated here.}

	\textit{\textbf{\textcolor{red}{Large-scale Datasets.}}} \textcolor{red}{As shown in the figure~\ref{fig: range big dataset}, On the 100M dataset, the top 3 ranked algorithms (WinFilter, UNIFY, and iRange) remain unchanged. However, their high performance is achieved at the cost of extremely high memory consumption, as detailed in Figure~\ref{fig:rangeFilter_search_memory_mb}. Meanwhile, VBASE and Faiss demonstrate moderate performance but significantly lower memory consumption than the top three algorithms}.

	\textit{\textbf{\textcolor{blue}{Multi-modal Queries}}}
	\textcolor{blue}{
	Similar to attribute filtering, the performance of multimodal queries with range filtering shown in Figure~\ref{fig:range-multimodal} is also consistent with the single-modality results in Figure~\ref{fig:exp_8_2}. As shown in Figure~\ref{fig:exp_8_3}, the algorithms still perform the worst on multimodal datasets like Text2Image. This is because these range query algorithms are almost graph-based. Graph algorithms need more iterations to find true neighbors in OOD datasets.}

%	\textit{\textbf{Multimodel, 85 and Deep100M.}} \textcolor{red}{The fig is Figure \ref{fig:TrinityExtension——Range}}
	\section{DISCUSSION}
	Based on the performance of the algorithms on different experimental scenarios, we discuss our findings as follows.

>>>>

	\noindent\textbf{\underline{Graph.}}
	Graph structures accelerate ANN search by connecting each data point with its nearest neighbors, as shown in Figure~\ref{fig:graph}. During querying, algorithms start from an initial point and use a greedy strategy to progressively move towards closer neighbors until convergence (e.g., the path from $c$ to $j$ in the 1NN case).

	Graph indexes are widely used in Hybrid ANN due to their efficiency. To support hybrid queries, in attribute filtering, a common practice involves attaching attribute metadata to nodes or edges, thereby skipping parts that do not satisfy query conditions during traversal (e.g., NHQ, Filtered-DiskANN, ACORN). In range filtering, edges typically store valid attribute intervals, and traversal proceeds only along edges that satisfy range constraints, thus narrowing the search space (e.g., SeRF, DSG).

%	Graph-based ANN search algorithms accelerate queries by building graph indexes on datasets. As illustrated in Figure~\ref{fig:graph}, each data point in the dataset corresponds to a point in the graph, and neighboring vertices (e.g., $a$ and $b$) are connected via edges based on their distance $\textit{dist}(a, b)$. In the graph index, each point maintains connections only to its nearest neighbors to preserve query efficiency. For instance, in Figure~\ref{fig:graph}, the gray point $c$ is connected to 5 neighbors $(a, b, d, e, f)$ and can traverse to them via corresponding edges.
%	
%	Given a query point $q$, the graph-based ANN search typically follows a greedy search strategy to find the $k$ nearest neighbors. For illustration, consider the case where $k = 1$ (i.e., 1NN): The algorithm initiates the search from an entry point (the gray point $c$). If there exists a neighbor $n$ (the point $f$) of the current point such that $\textit{dist}(n, q) < \textit{dist}(c, q)$, the algorithm updates the current point to $n$. This process is iteratively repeated—always moving to the neighbor closest to $q$—until no neighbor is found that is closer than the current point. The final point reached (the blue point $j$) is returned as the approximate nearest neighbor of $q$.
%	
%	\textcolor{violet}{Graph-based indexes are widely adopted in Hybrid ANN search algorithms due to their excellent performance. To support hybrid queries, existing approaches usually augment the original graph structures by explicitly associating attribute or range information with vertices or edges. In attribute filtering, a common approach is to attach attribute metadata to vertices or edges, allowing the traversal process to skip nodes or edges that do not meet attribute conditions of the query (e.g., NHQ, Filtered-DiskANN, ACORN). In range filtering, edges typically store valid attribute intervals, and traversal proceeds only along edges satisfying the range constraints, effectively narrowing down the search space (e.g., SeRF, DSG).}

	\noindent\textbf{\underline{IVF.}} 
	IVF divides data into multiple clusters through clustering (e.g., K-Means), with each cluster represented by a center (Figure~\ref{fig:ivf}). During querying, algorithms search only within the few clusters closest to the query point, which reduces computational load.

	IVF-based indexes also offer unique advantages in Hybrid ANN search scenarios, including natural support for pre-filtering, low memory footprint, and small index size. Efficient filtering is achieved by excluding vectors that do not satisfy attribute constraints before distance computation (e.g., Faiss, Milvus). Furthermore, hierarchical partitioning by attributes can form finer-grained sub-clusters to enhance filtering efficiency (e.g., CAPS, PUCK).
%	IVF-based ANN search algorithms improve computational efficiency by partitioning the dataset into clusters and restricting the search to clusters nearest to the query point. Specifically, as shown in Figure~\ref{fig:ivf}, IVF first selects a subset of data points and applies a clustering algorithm (e.g., K-Means) to obtain a set of centroids $(C_0, C_1, ..., C_6)$, each representing a distinct cluster. Every data point (gray dots) is then assigned to the cluster corresponding to its nearest centroid based on the distance metric, thereby forming the index structure.
%	
%	During the query process, given a query point $q$, the algorithm computes distances between $q$ and all centroids, selects a small number of the closest centroids ($C_0, C_1, C_2$), and then performs a search only within the associated clusters to identify the approximate nearest neighbors.
%	
%	\textcolor{violet}{IVF-based indexes also offer unique advantages in Hybrid ANN search scenarios, such as naturally supporting pre-filtering and requiring significantly lower memory consumption and smaller index sizes compared to graph-based indexes. With minor modifications, IVF indexes achieve attribute and range filtering by directly excluding vectors that do not meet query constraints through pre-filtering, thus skipping irrelevant vectors during intra-cluster distance computations (e.g., Faiss, Milvus). Moreover, clusters can be further partitioned hierarchically based on attributes (hierarchical partitioning), resulting in finer-grained sub-partitions that significantly enhance the efficiency of attribute filtering (e.g., CAPS, PUCK).}

@@

	\textit{\textbf{\textcolor{blue}{Different Hardware Environments.}}} 
%	\textcolor{cyan}{To investigate the impact of different server configurations on the performance of the algorithms, we conducted additional experiments on Server B using the YT-Audio-Real dataset, and the results are shown in Figure~\ref{fig:Different Computer}. Compared with the experimental results on Server A shown in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}f, the relative ranking of the algorithms remains largely unchanged, indicating that the performance of these methods is not significantly affected by differences in CPU configurations. Due to the lower CPU frequency and smaller RAM capacity of Server B compared to Server A, the QPS values are slightly lower across all methods.}
	\textcolor{red}{To investigate the impact of different server configurations on the performance of the algorithms, we conduct additional experiments on a server equipped with dual Intel(R) Xeon(R) Gold 6330 CPUs @ 2.00GHz using the YT-Audio-Real dataset, and the results are shown in Figure~\ref{fig:Different Computer}. Compared with the experimental results on the original server (as shown in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}(f)), the relative ranking of the algorithms remains largely unchanged, indicating that the performance of these methods is not significantly affected by differences in CPU configurations. Due to the lower CPU frequency and smaller RAM capacity, the QPS values are slightly lower across all methods.}

	\textit{\textbf{\textcolor{blue}{Large-scale Datasets.}}} 
	\textcolor{violet}{As shown in the figure~\ref{fig:Big Dataset}, On the 100M dataset, UNG and NHQ still exhibit excellent performance, demonstrating their good adaptability to large-scale datasets; however, this also comes with significant memory overhead. The clustering-based algorithm Faiss+HQI\_Batch shows moderate performance but has very low memory overhead, which makes it highly attractive for practical applications. We did not present the experimental results for StitchedVamana because we found that its recall consistently remained at a very low level. Its search starting point is set to 0 by default, and when the entry point is far from the target area, it easily gets stuck in a local optimum, making it difficult to navigate to the vicinity of the correct point.}

@@

\textcolor{blue}{
	We conducted experiments on the text2image-mix dataset. This dataset contains both image and text vectors. It is a more complex multimodal dataset. As shown in Figure~\ref{fig:MultiModal}, although the queries are more complex, UNG and NHQ still perform the best. CAPS is the next best.
	Since the dataset mixes two modalities, the ground truth of the query set usually corresponds to the same modality. The nearest neighbors in the index are often from the same modality because they are closer in distance. Therefore, the search results on this dataset are similar to those in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}. UNG and NHQ perform the best. CAPS clusters based on vector similarity. It actually clusters the vectors into the same modality. Then it uses attribute filtering to divide the data further. This allows CAPS to quickly find nearest neighbors that meet the filter conditions.
	In contrast, the text2image dataset shown in Figure~\ref{fig:exp_6_1} is different. The queries and the dataset are from different modalities. The queries are far from the dataset. The ground truths are also scattered and not clustered together. To find the ground truths, the algorithm often needs to traverse more points. This is a greater challenge for the algorithms.
	%	We conducted extensive scalability experiments on the Text2Image-Mix dataset. This dataset contains both image and text vectors. It represents a more complex multimodal query scenario. As shown in Figure 12, the UNG and NHQ algorithms still achieve the best performance, followed closely by CAPS.
	%	In the Text2Image-Mix dataset, the ground truth of the query vector is usually the corresponding vector in the same modality. Since the neighbor relationships in the graph are often based on vector distance, vectors of similar modalities naturally form closer connections in the graph. Therefore, during the search process, the algorithm tends to traverse efficiently within the same modality. This makes the search mechanism similar to our experiments on single-modality datasets, thus maintaining high performance.
	%	The good performance of the CAPS algorithm is due to its unique hierarchical space partitioning strategy. It first clusters vectors based on similarity. In multimodal data, this clustering naturally groups vectors of similar modalities into the same cluster. Then, CAPS further partitions these clusters using attributes. This allows it to quickly find the nearest neighbors that meet the filtering conditions.
	%	It is worth noting that, compared to the OOD dataset discussed in the "Dataset Impact" section, the Text2Image-Mix dataset also involves modality differences. However, its complexity lies in modality mixing. The OOD dataset, on the other hand, has a modality gap. This often requires traversing a broader and more sparse search space to find the ground truth. It poses a more severe challenge to the algorithm.
a}

@@

	\textit{\textbf{\textcolor{blue}{Different Hardware Environments.}}}
	\textcolor{cyan}{Consistent with Section 4.2.4, we conduct additional range filtering experiments on the YT-Audio-Real dataset using the same server configuration, with results presented in Figure~\ref{fig:Different_Computer_range}. When compared with the experimental results of the original server shown in Figure~\ref{fig:exp_8_2}a, our findings remain consistent with those reported in Section 4.2.4. For brevity and due to space limitations, we omit further discussion of these redundant results.}
%	\textcolor{cyan}{Consistent with Section 4.2.4, we conducted additional range filtering experiments using the YT-Audio-Real dataset on the same server, with results shown in Figure \ref{fig:Different Computer range}. Compared with the original server's experimental results (Figure \ref{fig:exp_8_2}($e$)), the conclusions remain identical to those described in Section 4.2.4. Due to space constraints, these details are not reiterated here.}

	\textit{\textbf{\textcolor{red}{Large-scale Datasets.}}} \textcolor{red}{As shown in the figure~\ref{fig:Big_Dataset_range}, On the 100M dataset, the top 3 ranked algorithms (WinFilter, UNIFY, and iRange) remain unchanged. However, their high performance is achieved at the cost of extremely high memory consumption, as detailed in Figure~\ref{fig:rangeFilter_search_memory_mb}. Meanwhile, VBASE and Faiss demonstrate moderate performance but significantly lower memory consumption than the top three algorithms}.

	\textit{\textbf{\textcolor{blue}{Multi-modal Queries}}}
	\textcolor{blue}{
	Similar to attribute filtering, the performance of multimodal queries with range filtering shown in Figure~\ref{fig:MultiModal_range} is also consistent with the single-modality results in Figure~\ref{fig:exp_8_2}. As shown in Figure~\ref{fig:exp_8_3}, the algorithms still perform the worst on multimodal datasets like Text2Image. This is because these range query algorithms are almost graph-based. Graph algorithms need more iterations to find true neighbors in OOD datasets.}

%	\textit{\textbf{Multimodel, 85 and Deep100M.}} \textcolor{red}{The fig is Figure \ref{fig:TrinityExtension——Range}}

	% Legend
	\includegraphics[width=0.98\columnwidth]{figures/exp/range_legend.pdf}
	\vspace{0.2cm} % Added a little space after the legend

	% --- Subfigure (a) containing two plots ---
	\begin{subfigure}[t]{0.98\columnwidth}
		\centering
		% First plot for the shared subfigure
		\includegraphics[width=0.507\linewidth]{figures/exp/range_multimodel.pdf}
		\hfill % Pushes the two plots apart
		% Second plot for the shared subfigure
		\includegraphics[width=0.48\linewidth]{figures/exp/range_multimodel_1.pdf}

		% Shared caption for the two plots above
		\captionsetup{font=small}
		\caption{MultiModal}
		\label{fig:MultiModal_range} % A single label for the combined subfigure
	\end{subfigure}

	% --- Subfigure (b) ---
	\begin{subfigure}[t]{0.507\columnwidth}
		\centering
		\includegraphics[width=\linewidth]{figures/exp/range_85.pdf}
		\captionsetup{font=small}
		\caption{Different Computer}
		\label{fig:Different_Computer_range}
	\end{subfigure}
	\hfill
	% --- Subfigure (c) ---
	\begin{subfigure}[t]{0.48\columnwidth}
		\centering
		\includegraphics[width=\linewidth]{figures/exp/range_deep100M.pdf}
		\captionsetup{font=small}
		\caption{Big Dataset}
		\label{fig:Big_Dataset_range}
	\end{subfigure}

	\caption{\textcolor{red}{TrinityExtension——Range}}
	\label{fig:TrinityExtension——Range}
\end{figure}

	\section{DISCUSSION}
	Based on the performance of the algorithms on different experimental scenarios, we discuss our findings as follows.
